Report on Late Winter2013 Production: Image Differencing\vspace{-2ex}

### 2013-03-29 W13report.tex
Background and positive vs. negative disussion.

We compared the measured numbers of false detections to the number we
expect due to random fluctuations in a background Gaussian field.  The
latter is a function of both seeing in the images, and of the
detection threshold used.  At the fiducial detection threshold of
5--sigma, the number of expected false positives per LSST sensor is of
order {\bf XX}, and we measure of order {\bf YY}, indicating that we
are able to achieve the systematic floor in false detection rate.  We
examined our measured false detection rate as a function of detection
threshold and find that it roughly scales as expected.  In order for
there to be less than 1 expected false detection per sensor, we must
operate at a detection threshold of {\bf X.X} sigma.

@@

The numbers of false detections in each best case are given in
Table~\ref{tab-bestfp10}.  We undertook a manual inspection of all
difference images to associate the {\tt DiaSources} with sources in the
images, and also use a 3'' association with the input source catalog
to define matches and orphans.  Associations are listed in the
N$_{Match}$ column of Table~\ref{tab-bestfp10}, with the unmatched
false positives in the N$_{Orphan}$ column.  The former are likely to
come from systematics in the subtractions of stars, while the latter
would be due to fluctuations in the background.  We note 2 effects
that may impact N$_{Match}$.  The first is that, with a 3'' matching
radius, 4\% of random fluctuations will by--chance associate with a
Source.  Assuming that N$_{Orphan}$ represents 96\% of the background
fluctuation population, we estimate the number of fluctuations from
that same population that would have associated with Sources as
N$_{Random}$.  Finally, we count all matches that occur within 100
pixels of the image boundary, where we might expect deficiencies in
the various spatial extrapolations (astrometry, backgrounds,
PSF--matching kernels) to yield systematic residuals, as N$_{D<100}$.
Figure~\ref{edgedist} provides the distribution of distances from the
edge of the sensor for all false detections in {\it blue}, with a
random sample of points in the image given in {\it green}.
Uncertainties are plotted using the square root of the number of
points in each bin, and the histograms are normalized to provide a
probability density.  Aside from the deconvolution reduction, we find
no significant correlation of false detections with the boundaries of
the fitting functions.  To the extent that N$_{Random}$ $<$
N$_{Match}$, we have a small residual set of systematic false
detections in our imaging ensemble.  Aside from the deconvolution run,
this is at the level of 10 total false detections over 9 sensors in 5
production runs, or 1 every 4--5 images. {\bf REDO BASED ON SIMON'S REANALYSIS}

We also list the ratio of negatively detected to positively detected
{\tt DiaSources}.  That this ratio is significantly greater than 1.0 in most
cases suggests that there is an over-subtraction of the background in
the process, which biases this ratio (see Section~\ref{sec-bg}).  For
completeness, the optimal configurations yielding these false detection
rates are given in Table~\ref{tab-bestconfig10}.

@@

\begin{figure}
\centering
\centering{\includegraphics[width=1.0\textwidth]{figures/fp_v_dsky.eps}}
\caption{
Misestimation of the sky value has an effect on both the number of false positives and relative abundance of positive false positives to negative false positives.
The top pane of this figure shows the relative number of +ve/-ve (-ve/+ve) if the sky background is under (over) subtracted at a particular fractional level.  The 
bottom panel shows how the total number of false positives changes for the same misestimation.  This is not a strong function of the seeing, so only the best
seeing example is shown.
}
\label{fig-skythresh}
\end{figure}

>>>>

We compared the measured numbers of false detections to the number we
expect due to random fluctuations in a background Gaussian field.  The
latter is a function of both seeing in the images, and of the
detection threshold used.  At the fiducial detection threshold of
5--sigma, the number of expected false positives per LSST sensor is of
order 3--10, and we measure a similar number, indicating that we are
able to achieve the systematic floor in false detection rate.  We
examined our measured false detection rate as a function of detection
threshold and find that it roughly scales as expected.  In order for
there to be less than 1 expected false detection per sensor (to
seeings as good as 0.6''), we must operate at a detection threshold of
5.5--sigma.

@@

The numbers N$_{Measured}$ of false detections in each best case are
given in Table~\ref{tab-bestfp10}, with the total number expected from
the Monte Carlo analysis given as N$_{Expect}$.  We undertook a manual
inspection of all difference images to associate the {\tt DiaSources}
with sources in the images, and also use a 3'' association with the
input source catalog to define matches and orphans.  Associations are
listed in the N$_{Match}$ column of Table~\ref{tab-bestfp10}, with the
unmatched false positives in the N$_{Orphan}$ column.  The former are
likely to come from systematics in the subtractions of stars, while
the latter would be due to fluctuations in the background.  We note 2
effects that may impact N$_{Match}$.  The first is that, with a 3''
matching radius, 4\% of random fluctuations will by--chance associate
with a Source.  Assuming that N$_{Orphan}$ represents 96\% of the
background fluctuation population, we estimate the number of
fluctuations from that same population that would have associated with
Sources as N$_{Random}$.  Finally, we count all matches that occur
within 100 pixels of the image boundary, where we might expect
deficiencies in the various spatial extrapolations (astrometry,
backgrounds, PSF--matching kernels) to yield systematic residuals, as
N$_{D<100}$.  Figure~\ref{edgedist} provides the distribution of
distances from the edge of the sensor for all false detections in {\it
  blue}, with a random sample of points in the image given in {\it
  green}.  Uncertainties are plotted using the square root of the
number of points in each bin, and the histograms are normalized to
provide a probability density.  Aside from the deconvolution
reduction, we find no significant correlation of false detections with
the boundaries of the fitting functions.  We note that we are
performing better than theory for visit v6866601 and prefiltering,
although this is only 1.4 times the scatter in the expectation value.
Otherwise, the results are as--expected from our simulations,
indicating that we are near the systematic floor in false detection
rate.  We also note that there is a slightly higher number of
associations with sources than we would expect randomly (N$_{Random}$
$<$ N$_{Match}$), although this is very small and at the rate of 1
every 4--5 images.

In the last column of Table~\ref{tab-bestfp10} we list the ratio of
negatively detected to positively detected {\tt DiaSources}.  This
ratio is significantly higher than 1.0 for all prefiltering visits,
suggesting that there is an over--subtraction of the background in the
image processing, which will bias this ratio (see
Section~\ref{sec-bg}).  The observed ratios of 2--3 suggest that there
is a background overestimation of order $1-2\%$ causing lower--sigma
fluctuations to be measured at 5--sigma (Figure~\ref{fig-skythresh}).
The postfiltering ratios vary by a similar amplitude, but with a
mixture of possible over-- and under--subtractions.

For completeness, the optimal configurations yielding these false
detection rates are given in Table~\ref{tab-bestconfig10}.

@@

\begin{figure}
\centering
\centering{\includegraphics[width=1.0\textwidth]{figures/fp_v_dsky.eps}}
\caption{
Misestimation of the sky value has an effect on both the number of false positives and relative abundance of positive false positives to negative false positives.
The top pane of this figure shows the relative number of +ve/-ve (-ve/+ve) detections if the sky background is under (over) subtracted at a particular fractional level.  The 
bottom panel shows how the total number of false positives changes for the same misestimation.  This is not a strong function of the seeing, so only the best
seeing example is shown.
}
\label{fig-skythresh}
\end{figure}
